### 2015-09-15 compl.tex
Text improvements

%% Proof for completeness theorem +foreword
%%%\begin{forrsi}
\section{Proof of the completeness theorem}
\label{app:compl}

Let us repeat our theorem and introduce minimal sequences again:
\[\begin{array}{r@{}l}
\forall S,S':&\\
&((\forall G: (SG\ne\ud\wedge S'G\ne\ud)\implies SG=S'G)\\
&\wedge\\
&(\exists F: SF\ne\ud\wedge S'F\ne\ud\wedge SF=S'F)\\
&\implies S \eqifnbr S'),
\end{array}\]
where \(G\) and \(F\) refer to filesystems.
%%%\end{forrsi}
%%%%%%%%%%%%%%%%%
%%%\begin{notrsi*}
%%%\newcommand{\varsection}[1]{\subsubsection{#1}}%
%%%\end{notrsi*}
%%%\begin{forrsi*}
\newcommand{\varsection}[1]{\subsection{#1}}%
%%%\end{forrsi*}

In the proof, by \(F\), we
mean a filesystem which satisfies the second condition. By \(G\), we
refer to any filesystem which satisfies \(SG\ne\ud\wedge S'G\ne\ud\).

Now consider the set of sequences \(\wp_S = \{ S^* | S \eqexp S^*\}\).
Because of our preconditions, the sequence \(break\) is not in
\(\wp_S\) (if it was, that is, \(S\eqexp break\), \(SF=\ud\) would hold
since by definition \(SF=(break)F\) if \(SF\ne\ud\)).

Let \(S_0\) be (one of) the shortest sequence(s) in \(\wp_S\) and
similarly
\(S^\prime_0\) (one of) the shortest sequence(s) in \(\wp_{S^\prime}\). 
%
The following holds for these sequences:
\[(SG\ne\ud\wedge S'G\ne\ud\implies) \quad SG=S'G=S_0G=S'_0G\ne\ud.\]
\begin{notrsi}
\emph{Proof.} Since 
\(SG\ne\ud\wedge S\eqexp S_0\implies SG=S_0G\) and
\(S'G\ne\ud\wedge S'\eqexp S'_0\implies S'G=S'_0G\); and since
\(SG\ne\ud\wedge S'G\ne\ud\) by assumption \(SG=S'G\ne\ud\) holds,
we have \(S_0G=SG=S'G=S'_0G\ne\ud\).

As a special case, we also know that \(SF=S'F=S_0F=S'_0F\ne\ud\).

\medskip
\end{notrsi}
Now let's investigate these ``minimal'' sequences.

@@

That way we also separated these commands from the other ones.
Now we focus on the remaining part: on the \(create\) and 
\(remove\) commands.

@@

\item[Lemma 3]
{\bf A command \(remove(\piesub)\) can not precede a command
\(create(\pi)\).} 
\begin{notrsi}
\emph{Proof.} This can be proved similarly, in this 
case the base step is
\(remove(\piesub); create({\pi})\) equals \(break\) or
expands to \(edit\) (Law \lawxxi~and \lawxxviii);
in the induction step we can prove that if
\(S[j-1]=create(\pisup)\) precedes \(S[j]=create(\pi)\), according to
the induction hypothesis, \(S\eqexp break\) 
(we know that \(\pisup\prec\pi\)). Otherwise the commands
would be able to be simplified (which contradicts the condition being
minimal sequence) or commuted (and therefore the induction hypothesis
still holds). (Laws \lawvii,
\lawviii, \lawxiv, \lawxv, \lawxx, \lawxxi, \lawxxviii.)
\end{notrsi}

@@

Now we prove that \(C\in S_0\Longleftrightarrow C\in S'_0\) for any
command \(C\). (Without loss of generality, it is enough to prove that 
\(C\in S_0\implies C\in S'_0\).)
\begin{itemize}
\item \(create(\pi,X)\in S_0\implies create(\pi,X)\in S'_0\). 
\emph{Proof.} By contradiction. First of all, we know that
\(F(\pi)=\ud\) otherwise \(S_0\) would break \(F\) and that would
contradict our assumption. Now assume that 
\(create(\pi,X)\not\in S'_0\). 
We have three cases. If there is no command on path \(\pi\) in
\(S'_0\) then \(S'_0F(\pi)=\ud\) and therefore 
\(S_0F(\pi)\ne S'_0F(\pi)\), that is, \(S_0F\ne S'_0F\) which contradicts
our assumption that \(S_0F=S'_0F\). If there was a command \(edit\) or
\(remove\) on path \(\pi\), it would break \(F\) since \(F(\pi)=\ud\) and
that again contradicts the precondition.

\item \(remove(\pi)\in S_0\implies remove(\pi)\in S'_0\).
\emph{Proof.} 
\begin{forrsi}
It can be done similarly to the case of \(create\).
\end{forrsi}
\begin{notrsi}
Now we can be sure that \(F(\pi)\ne\ud\) and
\(S_0F(\pi)=\ud\). If (instead of
\(remove\)) there was no command on \(\pi\) in \(S'_0\) then 
\(S'_0F(\pi)\ne\ud\)  would hold. In case of \(edit\), 
\(S'_0F(\pi)\ne\ud\) would hold (contradicting
\(S_0F(\pi)=\ud\)). \(create\) would break \(F\) since
\(F(\pi)\ne\ud\). Again, we have contradiction in all cases.
\end{notrsi}

\item \(edit(\pi,X)\in S_0\implies edit(\pi,X)\in S'_0\).
\emph{Proof.} Similarly we cannot have \(create\) or \(remove\) on path
\(\pi\) in sequence \(S'_0\) since \(F(\pi)\ne\ud\) and 
\(S'F(\pi)=SF(\pi)\ne\ud\). 
But now we might have no commands on
\(\pi\) in \(S'_0\): that way if \(F(\pi)=X\) then \(S_0F(\pi)=X\) and
\(S'_0F(\pi)=X\), too. 
But clearly \(\neg(S_0\eqifnbr S'_0)\). To prove that this is also
impossible, we need the first condition about all filesystems.

Define
\(H\) as \(F\{\pi\mapsto Y\}\) where \(Y\) has the same type
(file/directory) as \(F(\pi)\). If \(SF\ne\ud\) then \(SH\ne\ud\) since no
command breaks the filesystem because of the contents (not the type) of a
file or directory. Also, \(S'F\ne\ud\implies S'H\ne\ud\). Now, as a
special case of \(S_0G=S'_0G\) (note that 
\(SH\ne\ud\wedge S'H\ne\ud\) holds) we know
that \(S_0H=S'_0H\). But \(S_0H(\pi)=edit(\pi,X)H(\pi)=X\) and
\(S'_0H(\pi)=H(\pi)=Y\). We have a contradiction, therefore the lemma is
proved.
\end{itemize}

@@

\medskip
Now we will show that they can be made exactly the same by the
commuting algebraic laws.

We know that \(edit\)s can be moved to the ends of the sequences and
ordered. Let's suppose they are arranged this way.
Since we know that the two sequences contain the same
commands, these parts of the sequences is the same.
Therefore we can omit them in the discussion. Now we focus on
\(create\)s and \(remove\)s as we did above.

@@

We prove that the commands can be ordered the same way by induction on the
length of the sequences. If it
is \(1\), the problem can be solved easily. If the length of the sequences
are \(i>1\), we can choose a command from \(S_0\) which has no
preceding commands (i.e., a command which must precede it).
It is sure that it has no precedents in \(S'_0\). Now we
can move this command to the front of the sequences. Now the rest of
the sequences have length \(i-1\) therefore we can order them according to
the induction hypothesis.

We obtain \(S_0^*\) by applying this method to \(S_0\) and
\({S^\prime_0}^*\) from \(S^\prime_0\). Since they contain the same
commands, and the resulting order is the same,
\(S_0^*\) and
\({S^\prime_0}^*\) are \emph{exactly the same}. Now we have
\(S\eqexp S_0\eqexp S_0^*\equiv{S^\prime_0}^*\eqshr S^\prime_0\eqshr
S^\prime\), that is, \(S\eqifnbr S^\prime\). Our theorem is proved.
\begin{forrsi}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\end{forrsi}

\begin{CMNT}
It is also can be shown that 
\(S\equiv S'\). \emph{Proof.} We already know
that \(S\eqifnbr S'\). It means that we have a chain of sequences, 
\(S\eqexp S_1\eqexp S_2\ldots S'\) so that we gained each sequence by
applying a law to the preceding one. If \(S\not\equiv S'\), we used
somewhere an `\(E\)'--marked law, namely \(S_i\eqexp S_{i+1}\). It could 
be Law \lawiib, \lawxxiv, \lawxxvii~or \lawxxviii. [??...]
\end{CMNT}

>>>>

%% Proof for completeness theorem +foreword
\section{Proof of the completeness theorem}
\label{app:compl}

Let us repeat our theorem and introduce minimal sequences again:
\[\begin{array}{r@{}l}
\forall S,S':&\\
&(\forall G: (SG\ne\ud\wedge S'G\ne\ud)\implies SG=S'G)\\
&\wedge\\
&(\exists F: SF\ne\ud\wedge S'F\ne\ud)\\
&\implies S \eqifnbr S',
\end{array}\]
where \(G\) and \(F\) refer to filesystems.
\newcommand{\varsection}[1]{\subsection{#1}}%

Throughout the proof, 
% by \(F\), we
% mean a filesystem which satisfies the second condition. 
by \(G\), we
refer to any filesystem which satisfies \(SG\ne\ud\wedge S'G\ne\ud\).

Now consider the set of sequences \(\wp_S = \{ S^* | S \eqexp S^*\}\).
Because of our preconditions, the sequence \(break\) is not in
\(\wp_S\) (if it was, that is, \(S\eqexp break\), \(SG=\ud\) would hold
based on the soundness theorem).

Let \(S_0\) be (one of) the shortest sequence(s) in \(\wp_S\) and
similarly
\(S^\prime_0\) (one of) the shortest sequence(s) in \(\wp_{S^\prime}\). 
%
The following holds for these minimal sequences:
\[(SG\ne\ud\wedge S'G\ne\ud\implies) \quad SG=S'G=S_0G=S'_0G\ne\ud.\]
\begin{notrsi}
\emph{Proof.} Since 
\(SG\ne\ud\wedge S\eqexp S_0\implies SG=S_0G\) and
\(S'G\ne\ud\wedge S'\eqexp S'_0\implies S'G=S'_0G\); and since
\(SG\ne\ud\wedge S'G\ne\ud\) by assumption \(SG=S'G\ne\ud\) holds,
we have \(S_0G=SG=S'G=S'_0G\ne\ud\).

% As a special case, we also know that \(SF=S'F=S_0F=S'_0F\ne\ud\).

\medskip
\end{notrsi}
Now let us investigate these minimal sequences.

@@

This way we also separated these commands from the rest.
Now we can focus on the remaining part: on the \(create\) and 
\(remove\) commands.

@@

\item[Lemma 3]
{\bf A command \(remove(\piesub)\) can not precede a command
\(create(\pi)\).} 
\begin{notrsi}
\emph{Proof.} This can be proved similarly, in this 
case the base step is
\(remove(\piesub); create({\pi})\) equals \(break\) or
expands to \(edit\) (Law \lawxxi~and \lawxxviii);
in the induction step we can prove that if
\(S[j-1]=create(\pisup)\) precedes \(S[j]=create(\pi)\), according to
the induction hypothesis, \(S\eqexp break\) 
(we know that \(\pisup\prec\pi\)). Otherwise it would be possible to
shorten the sequence (which contradicts our precondition) or commuted (and therefore the induction hypothesis
still holds). (Laws \lawvii,
\lawviii, \lawxiv, \lawxv, \lawxx, \lawxxi, \lawxxviii.)
\end{notrsi}

@@

Now we prove that \(C\in S_0\Longleftrightarrow C\in S'_0\) for any
command \(C\). (Without loss of generality, it is enough to prove that 
\(C\in S_0\implies C\in S'_0\).)
\begin{itemize}
\item \(create(\pi,X)\in S_0\implies create(\pi,X)\in S'_0\). 
\emph{Proof.} By contradiction. First of all, we know that
\(G(\pi)=\ud\) otherwise \(S_0\) would break \(G\) and that would
contradict our assumption. Now assume that 
\(create(\pi,X)\not\in S'_0\). 
We have three cases. If there is no command on path \(\pi\) in
\(S'_0\) then \(S'_0G(\pi)=\ud\) and therefore 
\(S_0G(\pi)\ne S'_0G(\pi)\), that is, \(S_0G\ne S'_0G\) which contradicts
our assumption. If there was a command \(edit\) or
\(remove\) on path \(\pi\), it would break \(G\) since \(G(\pi)=\ud\) and
that again contradicts the precondition.

\item \(remove(\pi)\in S_0\implies remove(\pi)\in S'_0\).
\emph{Proof.} 
\begin{forrsi}
It can be done similarly to the case of \(create\).
\end{forrsi}
\begin{notrsi}
Now we can be sure that \(F(\pi)\ne\ud\) and
\(S_0G(\pi)=\ud\). If (instead of
\(remove\)) there was no command on \(\pi\) in \(S'_0\) then 
\(S'_0G(\pi)\ne\ud\)  would hold. In case of \(edit\), 
\(S'_0G(\pi)\ne\ud\) would hold (contradicting
\(S_0G(\pi)=\ud\)). \(create\) would break \(G\) since
\(G(\pi)\ne\ud\). Again, we have contradiction in all cases.
\end{notrsi}

\item \(edit(\pi,X)\in S_0\implies edit(\pi,X)\in S'_0\).
\emph{Proof.} Similarly we cannot have \(create\) or \(remove\) on path
\(\pi\) in sequence \(S'_0\) since \(G(\pi)\ne\ud\) and 
\(S'G(\pi)=SG(\pi)\ne\ud\). 
But now we might have no commands on
\(\pi\) in \(S'_0\): that way if \(G(\pi)=X\) then \(S_0G(\pi)=X\) and
\(S'_0G(\pi)=X\), too. 
% But clearly \(\neg(S_0\eqifnbr S'_0)\). 
To prove that this is also
impossible, we need the first condition about all filesystems.

Define
\(H\) as \(G\{\pi\mapsto Y\}\) where \(Y\) has the same type
(file/directory) as \(G(\pi)\). If \(SG\ne\ud\) then \(SH\ne\ud\) since no
command breaks the filesystem because of the contents (not the type) of a
file or directory. Also, \(S'G\ne\ud\implies S'H\ne\ud\). Now, 
from the coditions of the theorem
and \(SH\ne\ud\wedge S'H\ne\ud\) we know
that \(S_0H=S'_0H\). But \(S_0H(\pi)=edit(\pi,X)H(\pi)=X\) and
\(S'_0H(\pi)=H(\pi)=Y\). We have a contradiction, therefore the lemma is
proved.
\end{itemize}

@@

\medskip
Now we will show that they can be made exactly the same using the
commuting algebraic laws.

We know that \(edit\)s can be moved to the ends of the sequences and
ordered. Let us suppose they are arranged this way.
Since we know that the two sequences contain the same
commands, these parts of the sequences is the same.
Therefore we can omit them in the discussion. Now we focus on
\(create\)s and \(remove\)s as we did above.

@@

We prove that the commands can be ordered the same way by induction on the
length of the sequences. If it
is \(1\), the problem can be solved easily. If the length of the sequences
is \(i>1\), we can choose a command from \(S_0\) which has no
command that must precede it.
It is sure that it has no precedents in \(S'_0\). Now we
can move this command to the front of the sequences. Now the rest of
the sequences have length \(i-1\) therefore we can order them according to
the induction hypothesis.

We obtain \(S_0^*\) and \(S^\prime_0^*\) by applying this method to \(S_0\) and
\(S^\prime_0\). Since they contain the same
commands, and the resulting order is the same,
\(S_0^*\) and
\({S^\prime_0}^*\) are \emph{exactly the same}. Now we have
\(S\eqexp S_0\eqexp S_0^*\equiv{S^\prime_0}^*\eqshr S^\prime_0\eqshr
S^\prime\), that is, \(S\eqifnbr S^\prime\). Our theorem is proved.

### 2015-09-15 sound.tex
Text improvements

%% Proof for the soundness theorem + foreword
%%%\begin{forrsi}
\section{Proof for the soundness theorem}
\label{app:theosound}

Let us repeat our statement:
We prove that for every two sequences of commands 
\begin{gather*}
S\equiv S^* \implies \forall F: SF=S^*F,\\
S\eqexp S^*\wedge SF\ne\ud \implies \forall F: SF=S^*F.%\\
%S\eqifnbr S^*\wedge SF\ne\ud \wedge S^*F\ne\ud 
%\implies \forall F: SF=S^*F.
\end{gather*}
%%%\end{forrsi}

Both statement is true if our axioms and inference rules are sound since
we can only gain equivalent (or extending) sequences from axioms
or by using inference rules.

First, we prove the soundness of our inference rules. The proof for the
first two inference rules are quite similar. For the first one, we know
that the intended interpretation of \(S_1\equiv S_2\) is true and our
statement is that for every \(S,S'\), the interpretation of
\(S;S_1;S'\equiv S;S_2;S'\) holds. 
We know that for every filesystem
\(F\), \((S;S_1;S')F=S'(S_1(SF))\) and similarly
\(((S;S_2;S')F=S'(S_2(SF))\). We know that for every
\(G\), \(S_1G=S_2G\) holds, and therefore
as a special case \(S_1(SF)=S_2(SF)\). Thus
\(\forall F: (S;S_1;S')F=((S;S_2;S')F\). This is the interpretation of our
statement.

For the second law, we now that the interpretation of 
\(S_1\eqexp S_2\) is true and we show that in that
case \(S;S_1;S'\eqexp S;S_2;S'\) also true for filesystems. 
We know that for every \(F\) if \(S_1F\ne\ud\) then \(S_1F=S_2F\). If
\((S;S_1;S')F=\ud\) then the relation is clearly true (since it refers to
only the case if the filesystem is not broken). In the other case, if
\((S;S_1;S')F=S'(S_1(SF))\ne\ud\), we know that \(SF\ne\ud\) since
%%\remark{`defined' is not quite defined}
a filesystem cannot be made non--broken again by applying commands to it.
Now since \(S_1\eqexp S_2\) is true therefore either 
\(S_1(SF)=\ud\) and the relation is true similarly to the case above;
or \(S_1(SF)\ne\ud\) and therefore
\(S_1(SF)=S_2(SF)\ne\ud\) and \(S'(S_1(SF))=S'(S_2(SF))\ne\ud\).

The soundness proof of the third law is quite simple since if two
sequences act the same on \emph{all} filesystems then they act the same on
a subset of the filesystems, too.

It remains to show the soundness of each individual law,
but all the laws were derived from the definitions of the
commands. For precise proofs,
Appendix~\ref{app:soundness} shows an example.

>>>>

%% Proof for the soundness theorem + foreword
\section{Proof for the soundness theorem}
\label{app:theosound}

Let us repeat our statement:
We prove that for every two sequences of commands 
\begin{gather*}
S\equiv S^* \implies \forall F: SF=S^*F,\\
S\eqexp S^* \implies \forall F:(SF\ne\ud \implies SF=S^*F).
\end{gather*}

Both statements are true if our axioms and inference rules are sound,
beacuse
we can only generate equivalent (or extending) sequences from axioms
or by using inference rules.

First, we prove the soundness of our inference rules. The proof for the
first two inference rules are quite similar. For the first one, we know
that the intended interpretation of \(S_1\equiv S_2\) is true and our
statement is that for every \(S,S'\), the interpretation of
\(S;S_1;S'\equiv S;S_2;S'\) holds. 
We know that for every filesystem
\(F\), \((S;S_1;S')F=S'(S_1(SF))\) and similarly
\(((S;S_2;S')F=S'(S_2(SF))\). We know that for every
\(G\), \(S_1G=S_2G\) holds, and therefore
as a special case \(S_1(SF)=S_2(SF)\). Thus
\(\forall F: (S;S_1;S')F=(S;S_2;S')F\). This is the interpretation of our
statement.

For the second law, we now that the interpretation of 
\(S_1\eqexp S_2\) is true and we show that
\(S;S_1;S'\eqexp S;S_2;S'\) also true for filesystems. 
We know that for every \(F\) if \(S_1F\ne\ud\) then \(S_1F=S_2F\). If
\((S;S_1;S')F=\ud\) then the relation is clearly true (since it refers to
only the case if the filesystem is not broken). Otherwise, if
\((S;S_1;S')F=S'(S_1(SF))\ne\ud\), we know that \(SF\ne\ud\) since
a filesystem cannot be made non--broken again by applying commands to it.
Now since \(S_1\eqexp S_2\) is true therefore either 
\(S_1(SF)=\ud\) and the relation is true similarly to the case above;
or \(S_1(SF)\ne\ud\) and therefore
\(S_1(SF)=S_2(SF)\ne\ud\) and \(S'(S_1(SF))=S'(S_2(SF))\ne\ud\).

The soundness proof of the third law is quite simple as if two
sequences act the same on \emph{all} filesystems then they act the same on
a subset of the filesystems, too.

It remains to show the soundness of each individual law,
which can be verified based on the definitions of the commands from which
the laws have been derived.
Appendix~\ref{app:soundness} shows an example of the precise proofs.
